Visualization and analysis of probability distributions of large temporal data

### 2021-11-12 Rmd/05-conclusion.Rmd
non-temporal index added

Exploratory time series analysis entails numerous iterations of identifying and summarizing temporal dependencies. It is common practice to divide time into years, months, weeks, days, and so on in order to make inferences at both finer and coarser scales. In the literature, the formalization of these temporal deconstructions (granularities) is limited to linear time granularities such as hours, days, weeks, and months that respect the linear progression of time and are non-repeating. Cyclic granularities that are repeating in nature are useful for finding patterns in temporal data. They can be circular, quasi-circular, or aperiodic in nature. Hour of the day and day of the week are examples of circular granularities; the day of the month is an example of a quasi-circular granularity; and public holidays and school holidays are examples of aperiodic granularities. Additionally, time deconstructions can be based on a time hierarchy. Thus, single-order-up granularities such as second of minute or multiple-order-up granularities such as second of hour can be envisioned. The definitions and rules defined in the literature for linear granularities are insufficient for describing various types of cyclic granularities. Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal and non-temporal index. It also allows for the manipulation of single- and multiple-order-up time granularities via cyclic calendar algebra.

@@

@@

With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially. In this case, conditional distributions include not only temporal dependency but also variables and their dependencies. This adds to the already high-dimensional data structures that result from studying distributions. Current research on visualizing multivariate data must account for time's inherent characteristics in order to efficiently explore distributions of multivariate time series data.

@@

### Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data

>>>>

Exploratory time series analysis entails numerous iterations of identifying and summarizing temporal dependencies. It is common practice to divide time into years, months, weeks, days, and so on in order to make inferences at both finer and coarser scales. In the literature, the formalization of these temporal deconstructions (granularities) is limited to linear time granularities such as hours, days, weeks, and months that respect the linear progression of time and are non-repeating. Cyclic granularities that are repeating in nature are useful for finding patterns in temporal data. They can be circular, quasi-circular, or aperiodic in nature. Hour of the day and day of the week are examples of circular granularities; the day of the month is an example of a quasi-circular granularity; and public holidays and school holidays are examples of aperiodic granularities. Additionally, time deconstructions can be based on a time hierarchy. Thus, single-order-up granularities such as second of minute or multiple-order-up granularities such as second of hour can be envisioned. The definitions and rules defined in the literature for linear granularities are insufficient for describing various types of cyclic granularities. Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal index. It also allows for the manipulation of single- and multiple-order-up time granularities via cyclic calendar algebra. The approach is generalizable to non-temporal hierarchical granularities with an ordered index.

@@

### Computing aperiodic cyclic granularities

Aperiodic cyclic granularities capture repetitions of aperiodic linear granularities. All of these are recurring events, but with non-periodic patterns which can still be visualized to understand repetitive behavior. The computation of cyclic aperiodic granularities requires first computing aperiodic linear granularities. Aperiodic events may be created using R packages such as `almanac`(Vaughan, 2020) and `gs`(Laird-Smith, 2020). The functions in `gravitas` package need to integrate with them to support computation of aperiodic cyclic granularities.

### Clustering

Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data

@@

With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially. In this case, conditional distributions include not only temporal dependency but also variables and their dependencies. This adds to the already high-dimensional data structures that result from studying distributions. This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing multivariate time series data for one individual. Then a similar approach of unsupervised clustering can be used to group multiple time series over multiple time granularities and multiple variables. This is a similar approach followed in this thesis for handling univariate time series.

@@
